---
title: "DOHMH New York City Restaurant Inspection Results"
output: 
  html_document:
    css: styles.css
    toc: true
    toc_float: true
    theme: sandstone
    runtime: shiny
---

```{r include=FALSE}
library(shiny)
library(dplyr)
library(tidyverse)
library(kableExtra)
```


## Dataset Description


#### This dataset is derived from the [DOHMH New York City Restaurant Inspection Results Dataset](https://data.cityofnewyork.us/Health/DOHMH-New-York-City-Restaurant-Inspection-Results/43nn-pn8j/about_data) but has been specifically processed to include only restaurant data from Manhattan for this project. It contains sustained or pending violation citations from inspections conducted within three years prior to the most recent inspection date for establishments in active status. Each restaurant is uniquely identified by its CAMIS (record ID). Restaurants with no violations are represented by a single row and marked as "no violations" in the ACTION field. The dataset also includes uninspected new establishments and may contain missing or inconsistent values due to data entry or transfer errors.


## Data Dictionary

#### Since the [original data](https://data.cityofnewyork.us/Health/DOHMH-New-York-City-Restaurant-Inspection-Results/43nn-pn8j/about_data) exceeds 100MB, we are unable to upload it to GitHub. Therefore, we performed preliminary processing and retained only the areas where the Borough is Manhattan (Borough in which the entity (restaurant) is located:
##### • 1 = MANHATTAN
##### • 2 = BRONX
##### • 3 = BROOKLYN
##### • 4 = QUEENS
##### • 5 = STATEN ISLAND
##### • Missing; 

#### NOTE: There may be discrepancies between the zip code and listed borough due to differences in an establishment's mailing address and physical location). The cleaned dataset can be downloaded [here](https://github.com/katherinewang0903/final_nyc_restaurant/blob/main/Manhattan_Restaurant_Inspection_Results.csv.).

```{r include=FALSE}
library(readr)
manhattan_data <- read_csv("Manhattan_Restaurant_Inspection_Results.csv")

view (manhattan_data)
```

#### Based on the data where the Borough is Manhattan, there are 94,616 observations and includes the following variables:

* `CAMIS`: Unique identifier for each restaurant establishment.
* `DBA`: Doing Business As (DBA) name of the establishment.
* `BORO`: Borough where the establishment is located (Cleaned data we used only includes Manhattan).
* `BUILDING`: Building number where the establishment is located.
* `STREET`: Street name where the establishment is located.
* `ZIPCODE`: ZIP Code of the establishment.
* `PHONE`: Phone number of the establishment.
* `CUISINE DESCRIPTION`: Type of cuisine served by the establishment.
* `INSPECTION DATE`: Date of the most recent inspection.
* `ACTION`: Action taken based on the inspection results.
* `VIOLATION CODE`: Code for the specific health code violation.
* `VIOLATION DESCRIPTION`: Description of the health code violation.
* `CRITICAL FLAG`: Indicator for critical health violations.
* `SCORE`: Score assigned based on inspection results.
* `GRADE`: Grade assigned to the establishment (A, B, C, etc.).
* `GRADE DATE`: Date when the grade was assigned.
* `RECORD DATE`: Date when the record was created or updated.
* `INSPECTION TYPE`: Type of inspection conducted (initial, reinspection, etc.).
* `Latitude`: Latitude coordinate of the establishment.
* `Longitude`: Longitude coordinate of the establishment.
* `Community Board`: Community Board where the establishment is located.
* `Council District`: Council District where the establishment is located.
* `Census Tract`: Census Tract where the establishment is located.
* `BIN`: Building Identification Number.
* `BBL`: Borough-Block-Lot number.
* `NTA`: Neighborhood Tabulation Area of the establishment.
* `Location Point1`: Geographic location of the establishment.










