---
title: "ANALYSIS of Distribution"
output: 
  html_document:
    code_folding: hide
    always_allow_html: true
    toc: true
    toc_float: true
    theme: cosmo
---
```{r setup, include = FALSE}
library(dplyr)
library(purrr)
library(broom)
library(readr)
library(janitor)
library(knitr)
library(ggplot2)
library(stringr)
library(viridis)
library(plotly)
```

## Grade Distribution Over Time

This section explores how restaurant grades have changed over time, with a focus on neighborhoods and cuisine types. Using interactive visualizations, we analyze the trends to uncover patterns in hygiene ratings across Manhattan. We also highlight Italian restaurants as a case study for evaluating trends in a specific cuisine.

```{r, message=FALSE, warning=FALSE}
data <- read_csv("Manhattan_Restaurant_Inspection_Results.csv") %>%
  clean_names()

if (!all(c("grade", "inspection_date") %in% colnames(data))) {
  stop("The required columns 'grade' and 'inspection_date' are not found in the dataset.")
}

time_data <- data %>%
  filter(!is.na(grade), !is.na(inspection_date)) %>%
  mutate(year = as.numeric(format(as.Date(inspection_date, format = "%m/%d/%Y"), "%Y")))

grade_trends_borough <- time_data %>%
  group_by(year, boro, grade) %>%
  summarise(count = n(), .groups = "drop")

viridis_colors <- viridis(n = length(unique(grade_trends_borough$grade)))

plot_borough <- plot_ly(
  grade_trends_borough,
  x = ~year,
  y = ~count,
  color = ~grade,
  colors = viridis_colors,
  type = "scatter",
  mode = "lines+markers",
  split = ~grade
) %>%
  layout(
    title = list(text = "Trends in Manhattan Restaurants Hygiene Grades Over Time"),
    xaxis = list(title = "Year"),
    yaxis = list(title = "Number of Restaurants"),
    legend = list(title = list(text = "Grade"), x = 1.1),
    plot_bgcolor = 'rgba(0, 0, 0, 0)',
    paper_bgcolor = 'rgba(0, 0, 0, 0)'
  )

grade_trends_cuisine <- time_data %>%
  filter(cuisine_description == "Italian") %>%
  group_by(year, cuisine_description, grade) %>%
  summarise(count = n(), .groups = "drop")

plot_cuisine <- plot_ly(
  grade_trends_cuisine,
  x = ~year,
  y = ~count,
  color = ~grade,
  colors = viridis_colors,
  type = "scatter",
  mode = "lines+markers",
  split = ~grade
) %>%
  layout(
    title = list(text = "Grade Distribution Over Time for Italian Restaurants in Manhattan"),
    xaxis = list(title = "Year"),
    yaxis = list(title = "Number of Restaurants"),
    legend = list(title = list(text = "Grade"), x = 1.1),
    plot_bgcolor = 'rgba(0, 0, 0, 0)',
    paper_bgcolor = 'rgba(0, 0, 0, 0)'
  )

plot_borough
plot_cuisine
```

## Regression Model
Regression Model Analysis
To better understand the factors influencing hygiene grades, we build a regression model. This includes:

Converting grades (A, B, C) to numerical values.
Categorizing Manhattan into regions (Downtown, Midtown, Uptown, Other) based on zip codes.
Cleaning violation codes for more reliable analysis.
<br>
Key Insights:
Purpose of QQ Plot: Assess how well residuals align with a normal distribution, validating the model's fit.
Interpretation: Points closely aligning with the blue line indicate a strong model fit. Deviations highlight areas for improvement.
```{r, message=FALSE, warning=FALSE}
manhattan_data <- read_csv("Manhattan_Restaurant_Inspection_Results.csv", na = c("NA", "", "."))

cleaned_data <- manhattan_data %>%
  janitor::clean_names() %>%  
  filter(
    !is.na(dba),                         
    !is.na(cuisine_description),       
    !is.na(grade),                       
    !is.na(zipcode),
    grade %in% c("A", "B", "C")
  ) %>%
  mutate(
    grade_numeric = case_when(
      grade == "A" ~ 1,
      grade == "B" ~ 2,
      grade == "C" ~ 3
    ),  
    cuisine_description = factor(cuisine_description),  
    inspection_type = factor(inspection_type),  
    violation_code = str_trim(as.character(violation_code)),  
    location = case_when(
      zipcode >= 10000 & zipcode <= 10025 ~ "Downtown",
      zipcode >= 10026 & zipcode <= 10040 ~ "Midtown",
      zipcode >= 10041 & zipcode <= 10282 ~ "Uptown",
      TRUE ~ "Other"
    )
  ) %>%
  mutate(location = factor(location)) 


cleaned_data <- cleaned_data %>%
  filter(grepl("^[A-Za-z0-9]+$", violation_code)) 


predictors <- c("cuisine_description", "inspection_type", "violation_code", "location")


univariable_results <- lapply(predictors, function(predictor) {
  
  formula <- as.formula(paste("grade_numeric ~", predictor))
  
  
  model <- glm(formula, data = cleaned_data, family = quasipoisson) 
  
  
  tidy(model) %>%
    mutate(predictor = predictor) %>%
    select(predictor, term, estimate, std.error, p.value)
})


results_table <- bind_rows(univariable_results) %>%
  filter(term != "(Intercept)")

kable(
  head(results_table, 10), 
  format = "markdown",
  col.names = c("Predictor", "Term", "Estimate", "Std. Error", "p-value"),
  caption = "Top 10 Univariable Ordinal Regression Results"
)

model_example <- glm(grade_numeric ~ cuisine_description, data = cleaned_data, family = quasipoisson)
residuals_data <- augment(model_example)
```


```{r, warning=FALSE}
ggplot(residuals_data, aes(sample = .resid)) +
  stat_qq() +
  stat_qq_line(color = "blue") +
  labs(
    title = "QQ Plot of Residuals",
    x = "Theoretical Quantiles",
    y = "Sample Quantiles"
  ) +
  theme_minimal()
```
In order to obtain this model, we first load the dataset "Manhathathattan_Tetaurant_Inspection-Results.csv" and filter out irrelevant or missing data.
And perform conversion:
Grade_ numeric: Convert grades (A, B, C) into numerical values (1, 2, 3) for regression analysis.
Location: The postal code is divided into four areas: city center, central city, upper city, and others.
Violation-code: Cleaned up to ensure it only contains alphanumeric values, removing unexpected entries such as dates or special characters. And in order to make the visualization more visually appealing, I chose to only display the first ten lines of content.

1. Purpose of QQplot:
QQplot is a diagnostic tool used to compare the residual distribution of a regression model with the theoretical normal distribution. This helps evaluate the goodness of fit of the model.
2. Plot interpretation:
Blue line: represents the expected distribution under complete normality.
Point: represents the actual residual of the regression model.
If these points are closely aligned with the blue line and the residuals are approximately normally distributed, it indicates that the model fits well.
Deviation from this line, especially at the tail, indicates potential issues with normality or outliers in the data.
3. Observation results:
If these points closely follow this line, it can be confirmed that the model assumptions (such as the linearity and normality of residuals) are reasonably satisfied.

Go [Home](index.html)