---
title: "ANALYSIS of Location vs. Grade"

output: 
  html_document:
    code_folding: hide
    always_allow_html: true
    toc: true
    toc_float: true
    theme: cosmo
---

```{r}
knitr::opts_chunk$set(collapse = TRUE, 
                      message = FALSE)
```


```{r setup, message=FALSE}
library(tidyverse)
library(plotly)
library(ggplot2)
library(knitr)
library(readr)
library(dplyr)
library(tidyr)
library(scales)
library(vcd)
library(tidyverse)
library(htmltools)
```

In this chapter, we analyze the relationship between locations and grades. Our analysis focuses exclusively on data where the grades are `A`, `B`, or `C`, and excludes records with grades labeled as N (Not Yet Graded), Z (Grade Pending), or P (Grade Pending issued upon reopening after a closure). To ensure accuracy and relevance, we begin by filtering out unnecessary data from the dataset before proceeding with the analysis. The location is divided by `Financial District`, `Chinatown`, `West Village`, `Upper East Side`, `Upper West Side`, `Harlem`, `Washington Heights`, `East Village`, `Chelsea`, `Battery Park City`, `Bowery`, `SoHo`, `Gramercy Park`, `Flatiron District`, `Carnegie Hill`, `Tribeca`, `Governors Island`, `Midtown`, `Roosevelt Island`, `Other`.
```{r}
# The same data cleaning process in demographics part, also only consider data with grade A, B, C.
manhattan_data = read_csv("Manhattan_Restaurant_Inspection_Results.csv", na = c("NA", "", "."))
str (manhattan_data)

zip_grade = manhattan_data %>%
  janitor::clean_names() %>%  
  filter(
    !is.na(dba),                         
    !is.na(cuisine_description),       
    !is.na(grade),                       
    !is.na(score),                       
    !is.na(zipcode),
    grade %in% c("A", "B", "C")
  ) %>% 
  mutate(region_town = case_when(
    zipcode >= 10000 & zipcode <= 10025 ~ "Downtown",
    zipcode >= 10026 & zipcode <= 10040 ~ "Midtown",
    zipcode >= 10041 & zipcode <= 10282 ~ "Uptown",
    TRUE ~ "Other" # For ZIP codes like 11371, 12345, etc.
  )) %>%
  mutate(
    region = case_when(
      zipcode %in% c(10004, 10005, 10006, 10007, 10008, 10038, 10271, 10041) ~ "Financial District",
      zipcode %in% c(10013, 10002) ~ "Chinatown",
      zipcode %in% c(10014, 10011) ~ "West Village",
      zipcode %in% c(10021, 10028, 10065, 10075, 10128, 10029) ~ "Upper East Side",
      zipcode %in% c(10023, 10024, 10025, 10069, 10115, 10121, 10123) ~ "Upper West Side",
      zipcode %in% c(10026, 10027, 10030, 10031, 10035, 10037, 10039) ~ "Harlem",
      zipcode %in% c(10032, 10033, 10034, 10040) ~ "Washington Heights",
      zipcode %in% c(10003, 10009) ~ "East Village",
      zipcode %in% c(10001, 10011) ~ "Chelsea",
      zipcode %in% c(10280, 10281, 10282) ~ "Battery Park City",
      zipcode %in% c(10002) ~ "Bowery",
      zipcode %in% c(10012, 10013) ~ "SoHo",
      zipcode %in% c(10003, 10010) ~ "Gramercy Park",
      zipcode %in% c(10001, 10003, 10010, 10011, 10016) ~ "Flatiron District",
      zipcode %in% c(10028, 10128) ~ "Carnegie Hill",
      zipcode %in% c(10007, 10013) ~ "Tribeca",
      zipcode %in% c(10004) ~ "Governors Island",
      zipcode %in% c(10016, 10106, 10103, 10104, 10105, 10106, 10107, 10110, 10111, 10112, 10113, 10017, 10119, 10018, 10019, 10169, 10153, 10020, 10121, 10022, 10036, 10118, 10119, 10120, 10151, 10165, 1076, 10158) ~ "Midtown",
      zipcode %in% c(10044) ~ "Roosevelt Island",
      TRUE ~ "Other"
    )
  )
```
### Chi-Square Test

To investigate if where restaurants are located influence their inspection grades, we produce `Chi-Square test`. The result is presented in the `Contingency Table Heatmap` below. The `Chi-Square statistics` is 584.26, and `p-value is less than 0.05`, leading us to reject the null hypothesis that `the region and grade are independent`. This indicates a there a significant difference in the distribution of grades among regions at the 0.05 significance level. <br>
The `Midtown` region exhibits the highest concentration of Grade A restaurants, as indicated by the darkest shade of purple. Other regions like `Upper East Side` and `Upper West Side` also show substantial numbers of Grade A restaurants. `Grades B and C` are less frequent across all regions, with the light shading indicating their relatively low counts.

```{r}
# Create a contingency table of zip_code vs. grade
contingency_table <- table(zip_grade$region, zip_grade$grade)

# Perform Chi-Square Test
chi_test <- chisq.test(contingency_table)

# Convert the contingency table to a matrix
contingency_matrix <- as.matrix(contingency_table) 
contingency_matrix <- contingency_matrix[order(contingency_matrix[, "A"], decreasing = TRUE), ]

# Extract the residuals (contributions to the Chi-Square statistic)
residuals_matrix <- chi_test$residuals


# Extract chi-square test statistic and p-value
chi_stat <- round(chi_test$statistic, 2)
chi_pval <- format.pval(chi_test$p.value, digits = 3)

# Add the chi-square test result as annotations
plot_ly(
  x = colnames(contingency_matrix),
  y = rownames(contingency_matrix),
  z = contingency_matrix,
  type = "heatmap",
  colors = colorRamp(c("white", "purple"))
) %>%
  layout(
    title = "Contingency Table Heatmap with Chi-Square Result",
    xaxis = list(title = "Grade"),
    yaxis = list(title = "Region"),
    colorbar = list(title = "Count"),
    annotations = list(
      list(
        x = 2.1, 
        y = 3,
        text = paste("Chi-Square Statistic:", chi_stat, "<br>P-Value:", chi_pval),
        showarrow = FALSE,
        font = list(size = 14)
      )
    )
  )
```
### Cramér’s V test

To investigate how strong the association is between located and grades, we produce `Cramér’s V test`. The `Cramér’s V Value is  0.085`, which is less than 0.1 (Weak association), so that we conclude they have `weak association`. <br>
This conclusion is visualized in theree pie charts below of different regions that we can't tell large difference among them.
```{r}
# Calculate Cramér’s V
cramer_v <- assocstats(contingency_table)$cramer

# Print Cramér’s V
print(paste("Cramér's V:", round(cramer_v, 3)))
```


```{r}

```
Go [Home](index.html)